Real Estate

### 2023-05-09 src/real_estate.Rmd
outputting metric data to excel files

library(tidyverse)
library(janitor)
library(googlesheets4)
library(DT)

city_realEstate_data <- read_sheet("https://docs.google.com/spreadsheets/d/10n55Gt0rySN96OGTQ2_8PzeTcX66n26sOlZBiuiAeBU/edit#gid=808954990")

select_MSAs <- read_sheet("https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit?usp=sharing")

@@

```{r}
# aggregate yardi data into MSAs by cbsa
MSA_realEstate_data <- city_realEstate_data %>% 
  clean_names() %>% 
  rename(avg_rent2022 = x2022_rent) %>% 
   # calculate change in rent
  mutate(rent_change = (avg_rent2022 - average_rent_2021)/average_rent_2021 * 100) %>% 
  group_by(cbsa) %>% 
  # calculate average values for MSAs with more than one city
  summarize(avg_rent_2021 = mean(average_rent_2021, na.rm = TRUE),
            avg_rent_2022 = mean(avg_rent2022, na.rm = TRUE),
            avg_occupancy_21 = mean(percent_occupancy_2021, na.rm = TRUE),
            avg_rent_change = mean(rent_change, na.rm = TRUE))

@@

```{r}
# add MSA_median rent data
compliedMSA_realEstate<- select_MSAs %>% 
  left_join(MSA_realEstate_data) %>% 
    # calculate rent change
  mutate(med_rent_change = (median_rent2022 - median_rent2021)/median_rent2021 * 100) %>% 
  # calculate income to homeprice
  mutate(house_price_2021 = as.numeric(house_price_2021)) %>% 
  mutate(income_to_homeprice = median_income/house_price_2021) %>% 
  # calculate rent to income ratio
  # multiply monthly rent x 12 for annual rent!
  mutate(med_rent_to_income = (median_rent2021*12)/median_income)

@@

```{r}
# normalize columns into new data table
n_MSA_realEstate <- compliedMSA_realEstate %>% 
  # normalize population data
  mutate(n_pop_change = (pop_percent_change20_21 - min(pop_percent_change20_21))/(max(pop_percent_change20_21)-min(pop_percent_change20_21))) %>%
  # normalize employment growth
  mutate(n_employ_change = (avg_emp_YoY - min(avg_emp_YoY)) / (max(avg_emp_YoY) - min(avg_emp_YoY))) %>% 
  # normalize rent to income
  mutate(n_med_rent_to_income = (med_rent_to_income- min(med_rent_to_income)) / (max(med_rent_to_income) - min(med_rent_to_income))) %>% 
  # normalize income to home price
  mutate(n_med_rent_change = (med_rent_change - min(med_rent_change)) / (max(med_rent_change) - min(med_rent_change))) %>% 
  mutate(n_income_to_homeprice = (income_to_homeprice - min(income_to_homeprice)) / (max(income_to_homeprice) - min(income_to_homeprice))) %>% 
  mutate(n_avg_occupancy = (avg_occupancy_21 - min(avg_occupancy_21, na.rm = TRUE)) / (max(avg_occupancy_21, na.rm = TRUE) - min(avg_occupancy_21, na.rm = TRUE))) %>% 
  # normalize rent change
  mutate(n_avg_rent_change = (avg_rent_change - min(avg_rent_change, na.rm = TRUE)) / (max(avg_rent_change, na.rm = TRUE) - min(avg_rent_change, na.rm = TRUE)))

@@

for (i in 1:size[1]) {
  MSA_x_data <- c(n_MSA_realEstate$n_pop_change[i],
                  n_MSA_realEstate$n_employ_change[i],
                  n_MSA_realEstate$n_avg_occupancy[i],
                  n_MSA_realEstate$n_med_rent_change[i],
                  n_MSA_realEstate$n_med_rent_to_income_inverted[i], 
                  n_MSA_realEstate$n_income_to_homeprice_inverted[i]
                  )
  n_MSA_realEstate$real_estate_score[i] = weighted.mean(MSA_x_data, real_estate_weights)
}

@@

```{r}
# create downloadable data table
datatable(n_MSA_realEstate, extensions = 'Buttons', options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel', 'pdf', 'print')
  )
)
```

```{r}
#write result to google sheet
# gs4_create("MSA_real_estate_score")
sheet_write(n_MSA_realEstate, ss = "https://docs.google.com/spreadsheets/d/1Ssmb7pUd4NFZocVp9C75SRMOV6tAJCCt0TjrCYBB8pY/edit?usp=sharing", sheet = "n_MSA_realEstate")
```

>>>>

library(tidyverse)
library(janitor) # clean_names()
library(googlesheets4)
library(DT)
library(openxlsx)
library(here)
source(here("R/normalize.R"))

city_occupancy_data <- read_sheet("https://docs.google.com/spreadsheets/d/10n55Gt0rySN96OGTQ2_8PzeTcX66n26sOlZBiuiAeBU/edit#gid=808954990") %>% 
  clean_names() %>% 
  rename(occ_rate2021 = percent_occupancy_2021)

MSA_dempgraphic_data <- read_sheet("https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit?usp=sharing", sheet = "MSA_demographics") %>% 
  clean_names() %>%

  inner_join(city_occupancy_data, by = join_by(cbsa== cbsa, anchor_city==city))

@@

```{r}
# calculate metrics
real_estate_metrics<- MSA_dempgraphic_data %>% 
    # calculate rent change
  mutate(med_rent_change = (med_rent2022 - med_rent2021)/med_rent2021 * 100,
         house_price_2021 = as.numeric(house_price_2021),
         income_to_homeprice = med_income/house_price_2021,
         med_rent_to_income = (med_rent2021*12)/med_income) %>%  # multiply monthly rent x 12 for annual rent!
  select(!c(state.y, occ_rate2021, popestimate2021, med_rent2021, med_rent2022, house_price_2021, med_income)) %>% 
  rename(pop_change = pop_percent_change20_21)

write.xlsx(real_estate_metrics, here("data/intermediate/metric_data.xlsx"), sheetName = "real_estate_metrics")

#sheet_write(compliedMSA_realEstate, ss = "https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit#gid=426650684", sheet = "real_estate_metrics")

@@

# Normalize Metrics

```{r}
n_real_estate_metrics <- real_estate_metrics %>%
  mutate(n_pop_change = normalize(pop_change),
         n_employ_change = normalize(employ_change),
         n_income_homprice = 1- normalize(income_to_homeprice), # subtracted from 1 so higher values are more favorable
         n_rent_income = 1 - normalize(med_rent_to_income),
         n_rent_change = normalize(med_rent_change)

        )

@@

```{r}
# normalize columns into new data table
n_MSA_realEstate <- compliedMSA_realEstate %>% 
  # normalize population data
  mutate(n_pop_change = (pop_percent_change20_21 - min(pop_percent_change20_21))/(max(pop_percent_change20_21)-min(pop_percent_change20_21))) %>%
  # normalize employment growth
  mutate(n_employ_change = (avg_emp_YoY - min(avg_emp_YoY)) / (max(avg_emp_YoY) - min(avg_emp_YoY))) %>% 
  # normalize rent to income
  mutate(n_med_rent_to_income = (med_rent_to_income- min(med_rent_to_income)) / (max(med_rent_to_income) - min(med_rent_to_income))) %>% 
  # normalize income to home price
  mutate(n_med_rent_change = (med_rent_change - min(med_rent_change)) / (max(med_rent_change) - min(med_rent_change))) %>% 
  mutate(n_income_to_homeprice = (income_to_homeprice - min(income_to_homeprice)) / (max(income_to_homeprice) - min(income_to_homeprice))) %>% 
  mutate(n_avg_occupancy = (occ_rate2021 - min(occ_rate2021)) / (max(occ_rate2021) - min(occ_rate2021)), na.rm = FALSE)

@@

for (i in 1:size[1]) {
  MSA_x_data <- c(n_MSA_realEstate$n_pop_change[i],
                  n_MSA_realEstate$n_employ_change[i],
                  n_MSA_realEstate$n_avg_occupancy[i],
                  n_MSA_realEstate$n_med_rent_change[i],
                  n_MSA_realEstate$n_med_rent_to_income_inverted[i], 
                  n_MSA_realEstate$n_income_to_homeprice_inverted[i]
                  )
  n_MSA_realEstate$real_estate_score[i] = sum(MSA_x_data * real_estate_weights)
}

@@

```{r}
wt_real_estate_metrics <-n_MSA_realEstate %>%

   mutate(_score = apply(.[, 4:10], 1, sum))
  mutate(weighted_pop_change = 0.35*n_pop_change,
                  weighted_emp_change = 0.20*n_employ_change,
                  weighted_occ = 0.15 * n_avg_occupancy,
                  weighted_rent_change = 0.10*n_med_rent_change,
                  weighted_rent_income = 0.10 * n_med_rent_to_income_inverted, 
                  weighted_income_homeprice = 0.10 * n_income_to_homeprice_inverted) %>% 
  mutate(real_estate_score = weighted_pop_change + weighted_emp_change + weighted_occ + weighted_rent_change + weighted_rent_income + weighted_income_homeprice)

```{r}
#write result to google sheet
#gs4_create("anchor_real_estate_score")
sheet_write(n_MSA_realEstate, ss = "https://docs.google.com/spreadsheets/d/1Ssmb7pUd4NFZocVp9C75SRMOV6tAJCCt0TjrCYBB8pY/edit?usp=sharing", sheet = "anchor_realEstate")
```
